---
title: "re_lm02"
author: "Simon Rüegg"
date: "2025-05-04"
output: html_document
---

# Report exercise

```{r}
#Loading the libaries
library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)
library(caret)
library(recipes)
```

## 1: Characteristics of the two diffrent data sites Davos and Laegern

```{r}
# Loading the data
flux_davos <- read.csv("../data/FLX_CH-Dav_FLUXNET2015_FULLSET_DD_1997-2014_1-3.csv") #Data from Davos
flux_laegern <- read.csv("../data/FLX_CH-Lae_FLUXNET2015_FULLSET_DD_2004-2014_1-4.csv") #Data from Laegern
```

Information of for the diffrent sites can be found in a file, which tells the characteristics for each site.

```{r}
# Loading the file with the characteristics for the diffrent sites
siteinfo_flux <- load("../data/siteinfo_fluxnet2015(2).rda") #siteinfo_fluxnet2015 get loaded in the environment
siteinfo_fluxnet2015 |>
  filter(sitename %in% c("CH-Dav", "CH-Lae")) #just getting the characteristic for the site of Davos and Laegern.
```

From this file above we get the different characteristics from Dabos and Leagern. Both sides have an similiar latitude. The site of Laegern is a bit more north. There is a big diffrens with the elavation. The site of Davos is nearly 1000 Meters higher the the site of Laegern. Davos has an elavation of 1639 meters above sealevel, Lagern on of 689 Meter above sealevel. There is also a diffrence in the igbp_land_use und the plant_functional_type. While Davos is mostly an evergreen needleleaf forest, Laegern is a mixed forest with mostly decidous broadleaf trees.

## 2: Training the KNN-models

Wrangling Data

```{r}
# Splitting Data for Davos
flux_davos <- flux_davos |>  
  
  # select only the variables we are interested in
  dplyr::select(TIMESTAMP,
                GPP_NT_VUT_REF,    # the target
                ends_with("_QC"),  # quality control info
                ends_with("_F"),   # includes all all meteorological covariates
                -contains("JSB")   # weird useless variable
                ) |>

  # convert to a nice date object
  dplyr::mutate(TIMESTAMP = lubridate::ymd(TIMESTAMP)) |>

  # set all -9999 to NA
  dplyr::mutate(across(where(is.numeric), ~na_if(., -9999))) |> 

  
  # retain only data based on >=80% good-quality measurements
  # overwrite bad data with NA (not dropping rows)
  dplyr::mutate(GPP_NT_VUT_REF = ifelse(NEE_VUT_REF_QC < 0.8, NA, GPP_NT_VUT_REF),
                TA_F           = ifelse(TA_F_QC        < 0.8, NA, TA_F),
                SW_IN_F        = ifelse(SW_IN_F_QC     < 0.8, NA, SW_IN_F),
                LW_IN_F        = ifelse(LW_IN_F_QC     < 0.8, NA, LW_IN_F),
                VPD_F          = ifelse(VPD_F_QC       < 0.8, NA, VPD_F),
                PA_F           = ifelse(PA_F_QC        < 0.8, NA, PA_F),
                P_F            = ifelse(P_F_QC         < 0.8, NA, P_F),
                WS_F           = ifelse(WS_F_QC        < 0.8, NA, WS_F)) |> 

  # drop QC variables (no longer needed)
  dplyr::select(-ends_with("_QC"))



# Splitting Data for Laegern
flux_laegern <- flux_laegern|>  
  
  # select only the variables we are interested in
  dplyr::select(TIMESTAMP,
                GPP_NT_VUT_REF,    # the target
                ends_with("_QC"),  # quality control info
                ends_with("_F"),   # includes all all meteorological covariates
                -contains("JSB")   # weird useless variable
                ) |>

  # convert to a nice date object
  dplyr::mutate(TIMESTAMP = lubridate::ymd(TIMESTAMP)) |>

  # set all -9999 to NA
  dplyr::mutate(across(where(is.numeric), ~na_if(., -9999))) |> 

  
  # retain only data based on >=80% good-quality measurements
  # overwrite bad data with NA (not dropping rows)
  dplyr::mutate(GPP_NT_VUT_REF = ifelse(NEE_VUT_REF_QC < 0.8, NA, GPP_NT_VUT_REF),
                TA_F           = ifelse(TA_F_QC        < 0.8, NA, TA_F),
                SW_IN_F        = ifelse(SW_IN_F_QC     < 0.8, NA, SW_IN_F),
                LW_IN_F        = ifelse(LW_IN_F_QC     < 0.8, NA, LW_IN_F),
                VPD_F          = ifelse(VPD_F_QC       < 0.8, NA, VPD_F),
                PA_F           = ifelse(PA_F_QC        < 0.8, NA, PA_F),
                P_F            = ifelse(P_F_QC         < 0.8, NA, P_F),
                WS_F           = ifelse(WS_F_QC        < 0.8, NA, WS_F)) |> 

  # drop QC variables (no longer needed)
  dplyr::select(-ends_with("_QC"))

```

Checking data of completeness

```{r}
# Checking for Davos
visdat::vis_miss(
  flux_davos,
  cluster = FALSE, 
  warn_large_data = FALSE
  ) # alot of missing values in of the variable LW_IN_F

# Checking for Laegern
visdat::vis_miss(
  flux_laegern,
  cluster = FALSE, 
  warn_large_data = FALSE
  ) # For laegern there are ni values for P_F. Not using these variabe for the modelling and comparing
```

Wrangling Data

```{r}
# Eliminating the variables LW_IN_F and P_F for Davos
flux_davos <- flux_davos |>
  dplyr::select(-LW_IN_F) |>
  dplyr::select(-P_F)

# Eliminating the variables LW_IN_F and P_F for Laegern
flux_laegern <- flux_laegern |>
  dplyr::select(-LW_IN_F) |>
  dplyr::select(-P_F)

```

Combine Davos and Lagern

```{r}
flux_davos_laegern <- bind_rows(flux_davos, flux_laegern)
```

## 2: Splitting data and Training KNN models

Splitting the Data into training set and test set

```{r}
# Data splitting of Davos
set.seed(123)  # for reproducibility
split_davos <- rsample::initial_split(flux_davos, prop = 0.8, strata = "VPD_F")
fluxes_davos_train <- rsample::training(split_davos)
fluxes_davos_test <- rsample::testing(split_davos)

# Data splitting of Laegern
set.seed(123)  # for reproducibility
split_laegern <- rsample::initial_split(flux_laegern, prop = 0.8, strata = "VPD_F")
fluxes_laegern_train <- rsample::training(split_laegern)
fluxes_laegern_test <- rsample::testing(split_laegern)

# Data splitting of the combinaton Davos & Laegern
set.seed(123)  # for reproducibility
split_davos_laegern <- rsample::initial_split(flux_davos_laegern, prop = 0.8, strata = "VPD_F")
fluxes_davos_laegern_train <- rsample::training(split_davos_laegern)
fluxes_davos_laegern_test <- rsample::testing(split_davos_laegern)
```

Training KNN model for site Davos

```{r}
pp_davos <- recipes::recipe(GPP_NT_VUT_REF ~ SW_IN_F + VPD_F + TA_F, 
                      data = fluxes_davos_train) |> 
  recipes::step_center(recipes::all_numeric(), -recipes::all_outcomes()) |>
  recipes::step_scale(recipes::all_numeric(), -recipes::all_outcomes())

set.seed(1982)
mod_cv_davos <- caret::train(pp_davos, 
                       data = fluxes_davos_train |> drop_na(), 
                       method = "knn",
                       trControl = caret::trainControl(method = "cv", number = 10),
                       tuneGrid = data.frame(k = c(2, 5, 10, 15, 20, 25, 30, 35, 40, 60, 100)),
                       metric = "MAE")
print(mod_cv_davos)
```

Training KNN model for site of Laegern

```{r}
pp_laegern <- recipes::recipe(GPP_NT_VUT_REF ~ SW_IN_F + VPD_F + TA_F, 
                      data = fluxes_laegern_train) |> 
  recipes::step_center(recipes::all_numeric(), -recipes::all_outcomes()) |>
  recipes::step_scale(recipes::all_numeric(), -recipes::all_outcomes())

set.seed(1982)
mod_cv_laegern <- caret::train(pp_laegern, 
                       data = fluxes_laegern_train |> drop_na(), 
                       method = "knn",
                       trControl = caret::trainControl(method = "cv", number = 10),
                       tuneGrid = data.frame(k = c(2, 5, 10, 15, 20, 25, 30, 35, 40, 60, 100)),
                       metric = "MAE")
print(mod_cv_laegern)
```

Training KNN model for the combination of Davos and Laegern

```{r}
pp_davos_laegern <- recipes::recipe(GPP_NT_VUT_REF ~ SW_IN_F + VPD_F + TA_F, 
                      data = fluxes_davos_laegern_train) |> 
  recipes::step_center(recipes::all_numeric(), -recipes::all_outcomes()) |>
  recipes::step_scale(recipes::all_numeric(), -recipes::all_outcomes())

set.seed(1982)
mod_cv_davos_laegern <- caret::train(pp_davos_laegern, 
                       data = fluxes_davos_laegern_train |> drop_na(), 
                       method = "knn",
                       trControl = caret::trainControl(method = "cv", number = 10),
                       tuneGrid = data.frame(k = c(2, 5, 10, 15, 20, 25, 30, 35, 40, 60, 100)),
                       metric = "MAE")
print(mod_cv_davos_laegern)
```

Model trained with data from Davos

```{r}
# fit KNN model for Davos
k_davos <- mod_cv_davos$finalModel$k #optimal k for davos

mod_knn_davos <- caret::train(
  pp_davos, 
  data = fluxes_davos_train |> drop_na(), 
  method = "knn",
  trControl = caret::trainControl(method = "none"),
  tuneGrid = data.frame(k = k_davos),
  metric = "RMSE"
)

source("../R/eval_model_rmse_rsq.R")

# Evaluate model for Davos versus Davos, Laegern and the combination of Davos and Laegern
davos_davos <- eval_model_rmse_rsq(mod = mod_knn_davos, df_train = fluxes_davos_train, df_test = fluxes_davos_test) #Davos vs. Davos
davos_laegern <- eval_model_rmse_rsq(mod = mod_knn_davos, df_train = fluxes_davos_train, df_test = fluxes_laegern_test) #Davos vs. Laegern
davos_davoslaegern <- eval_model_rmse_rsq(mod = mod_knn_davos, df_train = fluxes_davos_laegern_train, df_test = fluxes_davos_laegern_test) #Davos vs. Davos-Laegern
```

Model trained with data from Laegern

```{r}
# fit KNN model for Davos
k_laegern <- mod_cv_laegern$finalModel$k #optimal k for davos

mod_knn_laegern <- caret::train(
  pp_laegern, 
  data = fluxes_laegern_train |> drop_na(), 
  method = "knn",
  trControl = caret::trainControl(method = "none"),
  tuneGrid = data.frame(k = k_laegern),
  metric = "RMSE"
)

source("../R/eval_model_rmse_rsq.R")

# Evaluate model for Davos versus Davos, Laegern and the combination of Davos and Laegern
laegern_laegern <- eval_model_rmse_rsq(mod = mod_knn_laegern, df_train = fluxes_laegern_train, df_test = fluxes_laegern_test) #Laegern vs. Laegern
laegern_davos <- eval_model_rmse_rsq(mod = mod_knn_laegern, df_train = fluxes_laegern_train, df_test = fluxes_davos_test) #Laegern vs. Davos
laegern_davoslaegern <- eval_model_rmse_rsq(mod = mod_knn_laegern, df_train = fluxes_laegern_train, df_test = fluxes_davos_laegern_test) # Laegern vs. Davos-Laegern
```

Model trained with data from the combination of Davos-Laegern

```{r}
# fit KNN model for Davos
k_davos_laegern <- mod_cv_davos_laegern$finalModel$k #optimal k for davos

mod_knn_davos_laegern <- caret::train(
  pp_davos_laegern, 
  data = fluxes_davos_laegern_train |> drop_na(), 
  method = "knn",
  trControl = caret::trainControl(method = "none"),
  tuneGrid = data.frame(k = k_laegern),
  metric = "RMSE"
)

source("../R/eval_model_rmse_rsq.R")

# Evaluate model for Davos versus Davos, Laegern and the combination of Davos and Laegern
davoslaegern_davoslaegern <- eval_model_rmse_rsq(mod = mod_knn_davos_laegern, df_train = fluxes_davos_laegern_train, df_test = fluxes_davos_laegern_test) #Laegern vs. Laegern
davoslaegern_davos <- eval_model_rmse_rsq(mod = mod_knn_davos_laegern, df_train = fluxes_davos_laegern_train, df_test = fluxes_davos_test) #Laegern vs. Davos
davoslaegern_laegern <- eval_model_rmse_rsq(mod = mod_knn_davos_laegern, df_train = fluxes_davos_laegern_train, df_test = fluxes_laegern_test) # Laegern vs. Davos-Laegern
```

## 3: Creating tables

Table for Davos

```{r}
table_davos <- data.frame(
  row.names = c("Evaluation against Laegern test set", "Evaluation agains Davos test set", "Evaluation against Laegern & Davos test set"),
  "RSQ" = c(davos_laegern[2], davos_davos[2], davos_davoslaegern[2]),
  "RMSE" = c(davos_laegern[1], davos_davos[1], davos_davoslaegern[1])
)

knitr::kable(table_davos, format = "html", caption = "Model trained on data from Davos") |>
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "center"
  )
```

Table for Laegern
```{r}
table_laegern <- data.frame(
  row.names = c("Evaluation against Laegern test set", "Evaluation agains Davos test set", "Evaluation against Laegern & Davos test set"),
  "RSQ" = c(laegern_laegern[2], laegern_davos[2], laegern_davoslaegern[2]),
  "RMSE" = c(laegern_laegern[1], laegern_davos[1], laegern_davoslaegern[1])
)

knitr::kable(table_laegern, format = "html", caption = "Model trained on data from Laegern") |>
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "center"
  )
```

Table for Davos & Laegern

```{r}
table_davoslaegern <- data.frame(
  row.names = c("Evaluation against Laegern test set", "Evaluation agains Davos test set", "Evaluation against Laegern & Davos test set"),
  "RSQ" = c(davoslaegern_laegern[2], davoslaegern_davos[2], davoslaegern_davoslaegern[2]),
  "RMSE" = c(davoslaegern_laegern[1], davoslaegern_davos[1], davoslaegern_davoslaegern[1])
)

knitr::kable(table_davoslaegern, format = "html", caption = "Model trained on data from Laegern & Davos") |>
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "center"
  )
```

## 4: Comparing of the different results

